Range Analysis of Whole Programs

### 2012-03-09 doc/Manuscript/paper.tex
Added section about experiments (layout only)
Added conclusion to the paper.

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
This paper describes an inter-procedural range analysis algorithm that scales
up to programs with millions of assembly instructions.
Contrary to many previous techniques, we handle comparisons between variables
without resorting to relational lattices or expensive algorithms.
We achieve path sensitiveness by using Bodik's Extended Static Single
Assignment form as the intermediate representation.
We show that by processing the strongly connected components that constitute
the graph of dependences between variables in topological order we not only
gain in time, but also in precision.
We have implemented our technique in LLVM, an industrial strenght compiler,
and have been able to process over 4 million assembly instructions in a
few seconds.
\end{abstract}

@@

@@

>>>>

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
This paper describes an inter-procedural range analysis algorithm that scales
up to programs with millions of assembly instructions.
Contrary to many previous techniques, we handle comparisons between variables
without resorting to relational lattices or expensive algorithms.
We achieve path sensitiveness by using Bodik's Extended Static Single
Assignment form as the intermediate representation.
We show that by processing the strongly connected components that constitute
the graph of dependences between variables in topological order we not only
gain in time, but also in precision.
We have implemented our technique in LLVM, an industrial strength compiler,
and have been able to process over 4 million assembly instructions in a
few seconds.
\end{abstract}

@@

% Absolute Time and complexity
% =============================

%Chart 1: shows complexity of the algorithm, plotting program size and time
% to run the analysis.

%Chart 2: time distribution for SPEC only

% Intra/Inter/Inline comparison
% =============================

% Chart 3: Time comparison between intra, inter and inlining:
% Each bar is inter/intra, or inline/intra
% We want to show that the whole program analysis is fast.

% (I) Chart 4: Precision comparison between intra/inter/inline
% *not for spec*: take five largest of bitwise, stanford and SPEC CINT.
% Add the three geomeans to the chart.

% (I) Chart 5: Same as (4), but this time precision in terms of [c, c], [-, c],
% [+, c], [-, +]. Perhaps take off [-, +]

% Impact of e-SSA form:
% =====================

% Chart 6: time overhead e-SSA. So, for each SPEC 2006, do time e-ssa/time ssa.

% (I) Chart 7: precision with e-SSA form.
% Percentage reduction (bitwidth only) for five largest of bitwise, stanford
% and SPEC CINT.

% Size overhead of different representations
% ==========================================

% Chart 8: Size overhead e-SSA/SSA. SPEC 2006 only.

% Chart 9: Size overhead inline/non-inline. SPEC 2006 only.

% Precision with different number of fixed iterations
% ===================================================

% Chart 10: Run inter + e-SSA with N*|SCC| in fixed update for five largest of
% bitwise, stanford and SPEC CINT.

% Table 1: Structural Properties for SPEC CPU 2006.
% Remember to add %SCC's with just one component.

% Impact of strongly connected components
% =======================================

@@

\section{Final Considerations}
\label{sec:con}

In this presentation we chose to omit correctness proofs for our algorithms.
For a proof that the widening and the narrowing operators give origin to
approximating sequences, we recommend Cousot and Cousot's work~\cite{Cousot}.
For a proof that the e-SSA form transformation is semantics preserving, we
invite the reader to check a recent report of ours~\cite{XXXX}.
In this work we also show that the results that we obtain via the
sparse analysis are equivalent to the results provided by a dense analysis.
In other words, if the dense analysis tells us that variable $v$ is
associated with the interval range $[l, u]$ at program point $i$, and $v'$
is the new name of $v$, alive at $i$ in the e-SSA form program, then $v'$ is
associated with the interval $[l, u]$ along its entire live range.

Our implementation of the range analysis algorithm described in this paper is
publicly available at \texttt{http://google.code/range-analysis/}.
This repository contains instructions about how to deploy and use our
implementation.
We provide a gallery of examples, including source codes,
CFGs and constraint graphs that we produce for meaningful programs at
\texttt{http://google.code/range-analysis/wiki/gallery}.
